Another Look at Regularized Discriminant Analysis for High-Dimensional Classification

### 2013-04-20 reports/paper-rda.Rnw
Updated the abstract and made some minor edits to the paper.

Microarrays have enabled researchers to measure gene-expression levels of
thousands of genes in a single experiment to identify a subset of genes and
their expression patterns that either characterize a particular cell state or
facilitate an accurate prediction of future cell state.  We focus on the
classification of tissues and other biological samples into one of $K$ disjoint,
predefined cancer types, where our goal is to ascertain the class membership of
a new tissue sample based on gene-expression microarray data. However,
microarray data typically consist of a large number of gene expressions $p$
relative to the total sample size $N$. Moreover, standard benchmark microarray
data sets usually consist of between $p = 10,000$ and $55,000$ probes and fewer
than $N = 100$ observations, in which case standard supervised-learning methods
exhibit poor classification performance because of the \emph{curse of
  dimensionality} \citep{Bellman:1961tn}. Furthermore, well-known classifiers,
including linear discriminant analysis \emph{(LDA)} and quadratic discriminant
analysis \emph{(QDA)}, are incalculable when $p > N$ because the sample
covariance matrices employed in both classifiers are singular. As a result, many
researchers, including \cite*{Mai:2012bf}, \cite*{Merchante:2012vk},
\cite*{Clemmensen:2011kr}, and \cite*{Huang:2010ju}, have introduced
restrictions on the population covariance matrices in the \emph{LDA} and
\emph{QDA} models to reduce number of covariance parameters that must be
estimated to ensure that the resulting classifiers are calculable.  For
instance, \cite{Dudoit:2002ev} have proposed the diagonal \emph{LDA}
(\emph{DLDA}) classifier, a modification of the \emph{LDA} classifier where the
class covariance matrices are assumed to be diagonal.

@@

In this paper we modify the \emph{RDA} classifier by first simplifying the
pooling matrix employed in the \emph{RDA} classifier. We construct a convex
combination of the covariance-matrix estimators employed in the \emph{LDA} and
\emph{QDA} classifiers and demonstrate that this simplification improves the
interpretation of the pooling operation of the \emph{RDA} classifier. The
\emph{LDA} classifier involves the assumption that the population covariance
matrices are equal, which is often too stringent in practice
\citep{Clemmensen:2011kr}. The \emph{HDRDA} classifier allows us to relax the
linearity assumption by including unequal covariance matrices in the model. The
\emph{HDRDA} classifier is then a compromise between the \emph{LDA} and
\emph{QDA} classifiers, so that the \emph{HDRDA} classifier includes both the
\emph{LDA} and \emph{QDA} classifiers as special cases. We then introduce a more
general shrinkage term to the \emph{RDA} classifier to establish our \emph{HDRDA}
classifier.

We show that the \emph{HDRDA} classifier is invariant to any adjustments of the
eigenvalues corresponding to the null space of the pooled sample covariance
matrix. Applying a reasoning similar to \cite{Ye:2006tq}, we decompose the
\emph{RDA} classifier and show that the sum related to the null space of the
pooled sample covariance matrix can be discarded. Hence, no gain in
classification performance can be obtained by adjusting these zero
eigenvalues. This invariance property addresses the computational shortcomings
of the \emph{RDA} classifier and provides a substantial increase in the
computational efficiency of the \emph{RDA} classifier and its model
selection. In fact, we show that the inverse and determinants can be reduced to
calculations of $n_k \times n_k$ matrices, where $n_k$ is the sample size of the
$k$th class, $k = 1, \ldots, K$. This reduction in calculation is clearly
advantageous because $n_k$ is often quite small, say, 15--20, in microarray data
sets.

In addition to the improvement in computational performance, we demonstrate that
our \emph{HDRDA} classifier is competitive with and often superior to several
recent classifiers designed for small-sample, high-dimensional data. In
particular, we compare our \emph{HDRDA} classifier with the \emph{DLDA}
classifier from \cite{Dudoit:2002ev} and the recently proposed classifiers from
\cite{Guo:2007te}, \cite{Pang:2009ik}, \cite{Tong:2012hw}, and
\cite{Witten:2011kc}. We conduct our comparison of the classification accuracy
of the competing classifiers using four small-sample, high-dimensional
microarray data sets.

@@

Here, we define our generalization of the \emph{RDA} classifier by first
simplifying the formulation of $\widehat{\bm \Sigma}_k(\lambda)$ in
\eqref{eq:sig-lambda} and demonstrating its clear interpretation as a linear
combination of the crossproducts of the training observations centered by their
respective class sample means. We define
\begin{align}
  \widehat{\bm\Sigma}_k(\lambda) = (1 - \lambda) \widehat{\bm\Sigma}_k + \lambda \widehat{\bm\Sigma}.\label{eq:sig-lambda-alternative}
\end{align}
as a convex combination of the MLEs of $\bm \Sigma_k$ and $\bm \Sigma$, where
$\lambda \in [0, 1]$ is the \emph{pooling} parameter. Thus, we avoid the convex
combination of $n_k$ and $N$ as defined by \cite{Friedman:1989tm}. By rewriting
\eqref{eq:sig-lambda-alternative} in terms of the observations $\bm x_i$, $i =
1, \ldots, N$, each centered by its class sample mean, we attain a clear
interpretation of $\widehat{\bm\Sigma}_k(\lambda)$. We have that
\begin{align}
	\widehat{\bm\Sigma}_k(\lambda) &= \widehat{\bm\Sigma}_k + \lambda (\widehat{\bm\Sigma} - \widehat{\bm\Sigma}_k) \nonumber\\
%	&= \widehat{\bm\Sigma}_k + \lambda \left(\sum_{k' = 1}^K \frac{n_k}{N} \widehat{\bm\Sigma}_{k'}  - \widehat{\bm\Sigma}_k \right) \nonumber\\
	&= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right) \widehat{\bm\Sigma}_k +  \frac{\lambda}{N} \sum_{\substack{k' = 1\\k' \ne k}} n_{k'} \widehat{\bm\Sigma}_{k'} \nonumber \\
%	&= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right)\sum_{i=1}^N I(y_i = k) \bm x_i \bm x_i' +  \frac{\lambda}{N} \sum_{i=1}^N I(y_i \ne k) \bm x_i \bm x_i' \nonumber \\
	&= \sum_{i=1}^N c_{ik}(\lambda) \bm x_i \bm x_i',\label{eq:sig-lambda-alternative2}
\end{align}
where $c_{ik}(\lambda) = \lambda N^{-1} + (1 - \lambda)n_k^{-1}I(y_i = k)$.
From \eqref{eq:sig-lambda-alternative2}, $\lambda$ weights the contribution of
each of the $N$ observations to estimating $\bm \Sigma_k$. That is, for $0 <
\lambda \le 1$, we estimate $\bm \Sigma_k$ with $N$ rather than $n_k$
observations. Below, we will show that the pooling operation is advantageous in
increasing the rank of each $\widehat{\bm\Sigma}_k(\lambda)$ from
rank$(\widehat{\bm\Sigma}_k)$ to rank$(\widehat{\bm\Sigma})$ for $0 < \lambda
\le 1$. Notice that if $\lambda = 0$, then the observations from the remaining
$K - 1$ classes do not contribute to the estimation of $\bm \Sigma_k$,
corresponding to $\widehat{\bm \Sigma}_k$. Furthermore, if $\lambda = 1$, the
weights in \eqref{eq:sig-lambda-alternative2} reduce to $1/N$, corresponding to
$\widehat{\bm\Sigma}$. For brevity, when $\lambda = 1$, we define $\bm X =
[\sqrt{c_{1k}(1)} \bm x_1', \ldots, \sqrt{c_{Nk}(1)} \bm x_N']'$ such that
$\widehat{\bm \Sigma} = \bm X' \bm X$. Similarly, for $\lambda = 0$, we define
$\bm X_k = [\sqrt{c_{1k}(0)} \bm x_1', \ldots, \sqrt{c_{Nk}(0)} \bm x_N']'$ such
that $\widehat{\bm \Sigma}_k = \bm X_k' \bm X_k$, $k = 1, \ldots, K$. In Figure
\ref{fig:hdrda-contours} we demonstrate the effect of $\lambda$ on the contours
of five multivariate normal populations with unequal covariance matrices.

@@

Next, we demonstrate that the \emph{HDRDA} classifier lends itself to an
efficient calculation of \eqref{eq:hdrda} and a rapid model selection of the
pooling parameter $\lambda$. We decompose \eqref{eq:hdrda} into a sum of two
components, where the first summand consists of matrix operations applied to
low-dimensional matrices and the second summand corresponds to the null space of
$\widehat{\bm \Sigma}$ in \eqref{eq:pooled-cov}. We show that the matrix
operations performed on the null space of $\widehat{\bm \Sigma}$ yield constant
quadratic forms across all classes and can be omitted under mild regularity
conditions. For $p \gg N$, the constant component involves determinants and
inverses of high-dimensional matrices, and by ignoring these calculations, a
substantial reduction in computational costs is achieved relative to Friedman's
\emph{RDA} classifier in \eqref{eq:rda}. Furthermore, a byproduct is that
adjustments to the associated eigenvalues have no effect on
\eqref{eq:hdrda}. Lastly, we exploit the so-called fast singular value
decomposition (SVD) to construct the eigenvalue decomposition of $\widehat{\bm
  \Sigma}$, further reducing the computational costs of our proposed \emph{RDA}
classifier.

@@

We now present our main result, where we decompose the \emph{HDRDA} decision rule
and show that the component requiring the largest computational costs does not
contribute to the classification of an unlabeled observation performed using
Lemma \ref{lemma:RDA-constant-term}. Our main result reduces \eqref{eq:hdrda} to
an equivalent, more computationally efficient and more stable decision rule.

\begin{thm}
Let $\tilde{\bm \Sigma}_k$ and $\bm W_k$ be defined as in \eqref{eq:rda-matrix}
and \eqref{eq:Wk}, respectively, and let $\bm U_1$ be defined as above. Then,
the \emph{HDRDA} decision rule in \eqref{eq:hdrda} is equivalent to
	\begin{align}
		D_{HDRDA}(\bm x) &= \argmin_k  (\bm x - \xbar_k)' \bm U_1 \bm W_k^{-1} \bm U_1' (\bm x - \xbar_k) + \ln | \bm W_k |. \label{eq:rda-decomposed}
	\end{align}

@@

Thus far we have shown that the \emph{HDRDA} classifier in
\eqref{eq:rda-decomposed} is invariant to the term $\bm U_2$, thus yielding a
substantial reduction in the computational costs. While the classifier is much
improved in terms of computational performance compared to the \emph{RDA}
classifier originally proposed by \cite{Friedman:1989tm}, we provide two
additional results that further improve the computational performance of our
\emph{HDRDA} classifier. First, we demonstrate that the inverse and determinants
of $\bm W_k$ can be written so that they are performed on $n_k \times n_k$
matrices rather than the larger $q \times q$ matrices, resulting in further
improvements to the computational efficiency of the \emph{HDRDA} classifier.

@@

In this section, we describe four high-dimensional microarray data sets and
compare our proposed \emph{HDRDA} classifier with five recently proposed
classifiers designed for small-sample, high-dimensional data: the \emph{DLDA}
classifier from \cite{Dudoit:2002ev}, the shrunken centroids \emph{RDA}
classifier from \cite{Guo:2007te}, the shrinkage-based \emph{DLDA} classifier
from \cite{Pang:2009ik}, the shrinkage-mean-based DLDA classifier from
\cite{Tong:2012hw}, and the penalized \emph{LDA} classifier from
\cite{Witten:2011kc}. Below, we refer to each classifier by the first author's
surname. We evaluated the classification performance of each classifier by
randomly partitioning the data such that $2/3$ of the observations were
allocated as training data and the remaining $1/3$ of the observations were
allocated as a test data set. For each random partition, we applied the
feature-selection method from \cite{Dudoit:2002ev} to choose the $d$ features
having the largest ratio of their between-group to within-group sums of squares
from the training data. We then filtered the test data set with the
corresponding $d$ features. After training each classifier from the filtered
training data, we classified the filtered test data sets and computed the
proportion of mislabeled test observations to estimate the classification error
rate for each classifier. Repeating this process 1000 times, we computed the
average of the error-rate estimates for each classifier. We remark that our
simulation design avoids the selection bias discussed in
\cite{Ambroise:2002fa}. We did not include the \emph{RDA} classifier from
\cite{Friedman:1989tm} in our studies to avoid its excessive computations. In
Table \ref{tab:data-summary}, we summarize the attributes of each microarray
data set.

@@

Next, we examine the optimal estimates of the tuning parameters for the
\emph{HDRDA} classifier. For brevity, we include only a graphical summary of the
optimal estimates of $\lambda$ for the Shipp data set in Figure
\ref{fig:optimal-shipp}. Similar figures for the other three data sets can be
found in the Supplementary Information. In Figure \ref{fig:optimal-shipp}, the
distribution of the optimal values of $\lambda$ selected through
cross-validation has a notable bathtub shape, resembling a Beta distribution
with both shape parameters less than unity. Specifically, the optimal values of
$\lambda$ are typically near 0 or near 1, corresponding to the \emph{QDA} and
\emph{LDA} classifiers, respectively. We noticed a similar distribution of
optimal values of $\lambda$ for the other three data sets.

@@

In our simulation study, we compared the \emph{HDRDA} classifier to several
recently proposed classifiers, including the Dudoit, Pang, Tong, and Guo
classifiers, each of which assumes independence of the features conditional on
class membership. Despite the rapid computational performance of diagonal
classifiers and the reduction in the number of parameters to estimate,
\cite{Mai:2012bf} have noted that diagonal classifiers can often yield inferior
classification performance to other classification methods. We have shown that
our proposed \emph{HDRDA} classifier often yields superior classification
accuracy to the diagonal classifiers considered, confirming the assertion of
\cite{Mai:2012bf}. Furthermore, the \emph{HDRDA} classifier is able to compete
with the diagonal classifiers in terms of speed by incorporating a new
dimension-reduction technique.

>>>>

Microarrays have enabled researchers to measure gene-expression levels of
thousands of genes in a single experiment to identify a subset of genes and
their expression patterns that either characterize a particular cell state or
facilitate an accurate prediction of future cell state.  We focus on the
classification of tissues and other biological samples into one of $K$ disjoint,
predefined cancer types, where our goal is to ascertain the class membership of
a new tissue sample based on gene-expression microarray data. However,
microarray data typically consist of a large number of gene expressions $p$
relative to the total sample size $N$. Moreover, standard benchmark microarray
data sets usually consist of between $p = 10,000$ and $55,000$ probes and fewer
than $N = 100$ observations, in which case standard supervised-learning methods
exhibit poor classification performance because of the \emph{curse of
  dimensionality} \citep{Bellman:1961tn}. Furthermore, well-known classifiers,
including linear discriminant analysis \emph{(LDA)} and quadratic discriminant
analysis \emph{(QDA)}, are incalculable when $p > N$ because the sample
covariance matrices employed in both classifiers are singular. As a result, many
researchers, including \cite*{Mai:2012bf}, \cite*{Fan:2012iq},
\cite*{Merchante:2012vk}, \cite*{Clemmensen:2011kr}, and \cite*{Huang:2010ju},
have introduced restrictions on the population covariance matrices in the
\emph{LDA} and \emph{QDA} models to reduce number of covariance parameters that
must be estimated to ensure that the resulting classifiers are calculable.  For
instance, \cite{Dudoit:2002ev} have proposed the diagonal \emph{LDA}
(\emph{DLDA}) classifier, a modification of the \emph{LDA} classifier where the
class covariance matrices are assumed to be diagonal.

@@

In this paper we modify the \emph{RDA} classifier by first simplifying the
pooling matrix employed in the \emph{RDA} classifier. We construct a convex
combination of the covariance-matrix estimators employed in the \emph{LDA} and
\emph{QDA} classifiers and demonstrate that this simplification improves the
interpretation of the pooling operation of the \emph{RDA} classifier.
\cite{Clemmensen:2011kr} have argued that the assumption of equal population
covariance matrices within the \emph{LDA} classifier is often too stringent in
practice. The \emph{HDRDA} classifier allows us to relax the linearity
assumption by including unequal covariance matrices in the model. The
\emph{HDRDA} classifier is then a compromise between the \emph{LDA} and
\emph{QDA} classifiers, so that the \emph{HDRDA} classifier includes both the
\emph{LDA} and \emph{QDA} classifiers as special cases. We then establish the
\emph{HDRDA} classifier by applying covariance-matrix regularization with a
parameterization that includes a large family of shrinkage methods, such as the
methods from \cite{Srivastava:2007ww}, \cite{Rao:1971ul}, and a wide variety of
other shrinkage methods studied by \cite{Ramey:2013ji} and \cite{Xu:2009fl}.

We show that the \emph{HDRDA} classifier is invariant to any adjustments of the
eigenvalues corresponding to the null space of the pooled sample covariance
matrix. Applying a reasoning similar to \cite{Ye:2006tq}, we decompose the
\emph{RDA} classifier and show that the sum related to the null space of the
pooled sample covariance matrix can be discarded. Hence, no gain in
classification performance can be obtained by adjusting these zero
eigenvalues. This invariance property addresses the computational shortcomings
of the \emph{RDA} classifier and provides a substantial increase in
computational efficiency and model selection. In fact, we show that the inverse
and determinants can be reduced to calculations of $n_k \times n_k$ matrices,
where $n_k$ is the sample size of the $k$th class, $k = 1, \ldots, K$. This
reduction in calculation is clearly advantageous because $n_k$ is often quite
small, say, 15--20, in microarray data sets.

We also demonstrate that our \emph{HDRDA} classifier is competitive with and
often superior to several recent classifiers designed for small-sample,
high-dimensional data. In particular, we compare our \emph{HDRDA} classifier
with the \emph{DLDA} classifier from \cite{Dudoit:2002ev} and the recently
proposed classifiers from \cite{Guo:2007te}, \cite{Pang:2009ik},
\cite{Tong:2012hw}, and \cite{Witten:2011kc}. We conduct our comparison of the
classification accuracy of the competing classifiers using four small-sample,
high-dimensional microarray data sets.

@@

Here, we define our adaptation of the \emph{RDA} classifier by first simplifying
the formulation of $\widehat{\bm \Sigma}_k(\lambda)$ in \eqref{eq:sig-lambda}
and demonstrating its clear interpretation as a linear combination of the
crossproducts of the training observations centered by their respective class
sample means. We define
\begin{align}
  \widehat{\bm\Sigma}_k(\lambda) = (1 - \lambda) \widehat{\bm\Sigma}_k + \lambda \widehat{\bm\Sigma}.\label{eq:sig-lambda-alternative}
\end{align}
as a convex combination of the MLEs of $\bm \Sigma_k$ and $\bm \Sigma$, where
$\lambda \in [0, 1]$ is the \emph{pooling} parameter. Thus, we avoid the convex
combination of $n_k$ and $N$ as defined by \cite{Friedman:1989tm}. By rewriting
\eqref{eq:sig-lambda-alternative} in terms of the observations $\bm x_i$, $i =
1, \ldots, N$, each centered by its class sample mean, we attain a clear
interpretation of $\widehat{\bm\Sigma}_k(\lambda)$. We have that
\begin{align}
	\widehat{\bm\Sigma}_k(\lambda) &= \widehat{\bm\Sigma}_k + \lambda (\widehat{\bm\Sigma} - \widehat{\bm\Sigma}_k) \nonumber\\
%	&= \widehat{\bm\Sigma}_k + \lambda \left(\sum_{k' = 1}^K \frac{n_k}{N} \widehat{\bm\Sigma}_{k'}  - \widehat{\bm\Sigma}_k \right) \nonumber\\
	&= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right) \widehat{\bm\Sigma}_k +  \frac{\lambda}{N} \sum_{\substack{k' = 1\\k' \ne k}} n_{k'} \widehat{\bm\Sigma}_{k'} \nonumber \\
%	&= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right)\sum_{i=1}^N I(y_i = k) \bm x_i \bm x_i' +  \frac{\lambda}{N} \sum_{i=1}^N I(y_i \ne k) \bm x_i \bm x_i' \nonumber \\
	&= \sum_{i=1}^N c_{ik}(\lambda) \bm x_i \bm x_i',\label{eq:sig-lambda-alternative2}
\end{align}
where $c_{ik}(\lambda) = \lambda N^{-1} + (1 - \lambda)n_k^{-1}I(y_i = k)$.
From \eqref{eq:sig-lambda-alternative2}, $\lambda$ weights the contribution of
each of the $N$ observations to estimate $\bm \Sigma_k$. That is, for $0 <
\lambda \le 1$, we estimate $\bm \Sigma_k$ with $N$ rather than $n_k$
observations. Below, we will show that the pooling operation is advantageous in
increasing the rank of each $\widehat{\bm\Sigma}_k(\lambda)$ from
rank$(\widehat{\bm\Sigma}_k)$ to rank$(\widehat{\bm\Sigma})$ for $0 < \lambda
\le 1$. Notice that if $\lambda = 0$, then the observations from the remaining
$K - 1$ classes do not contribute to the estimation of $\bm \Sigma_k$,
corresponding to $\widehat{\bm \Sigma}_k$. Furthermore, if $\lambda = 1$, the
weights in \eqref{eq:sig-lambda-alternative2} reduce to $1/N$, corresponding to
$\widehat{\bm\Sigma}$. For brevity, when $\lambda = 1$, we define $\bm X =
[\sqrt{c_{1k}(1)} \bm x_1', \ldots, \sqrt{c_{Nk}(1)} \bm x_N']'$ such that
$\widehat{\bm \Sigma} = \bm X' \bm X$. Similarly, for $\lambda = 0$, we define
$\bm X_k = [\sqrt{c_{1k}(0)} \bm x_1', \ldots, \sqrt{c_{Nk}(0)} \bm x_N']'$ such
that $\widehat{\bm \Sigma}_k = \bm X_k' \bm X_k$, $k = 1, \ldots, K$. In Figure
\ref{fig:hdrda-contours} we plot the contours of five multivariate normal
populations with unequal covariance matrices and $\lambda = 0$. As $\lambda$ is
increased towards 1, the contours become more similar, ultimately resulting in
equal covariance matrices and contours for the parsimonious case of $\lambda =
1$.

@@

Next, we demonstrate that the \emph{HDRDA} classifier lends itself to an
efficient calculation of \eqref{eq:hdrda} and a rapid model selection of the
pooling parameter $\lambda$. We decompose \eqref{eq:hdrda} into a sum of two
components, where the first summand consists of matrix operations applied to
low-dimensional matrices and the second summand corresponds to the null space of
$\widehat{\bm \Sigma}$ in \eqref{eq:pooled-cov}. We show that the matrix
operations performed on the null space of $\widehat{\bm \Sigma}$ yield constant
quadratic forms across all classes and can be omitted under mild regularity
conditions. For $p \gg N$, the constant component involves determinants and
inverses of high-dimensional matrices, and by ignoring these calculations, a
substantial reduction in computational costs is achieved relative to Friedman's
\emph{RDA} classifier in \eqref{eq:rda}. Furthermore, a byproduct is that
adjustments to the associated eigenvalues have no effect on
\eqref{eq:hdrda}. Lastly, we utilize the singular value decomposition (SVD) to
calculate efficiently the eigenvalue decomposition of $\widehat{\bm \Sigma}$,
further reducing the computational costs of our proposed \emph{RDA} classifier.

@@

We now present our main result, where we decompose the \emph{HDRDA} decision
rule and show that the component requiring the largest computational costs does
not contribute to the classification of an unlabeled observation performed using
Lemma \ref{lemma:RDA-constant-term}. Hence, we reduce \eqref{eq:hdrda} to an
equivalent, more computationally efficient decision rule.

\begin{thm}
Let $\tilde{\bm \Sigma}_k$ and $\bm W_k$ be defined as in \eqref{eq:rda-matrix}
and \eqref{eq:Wk}, respectively, and let $\bm U_1$ be defined as above. Then,
the \emph{HDRDA} decision rule in \eqref{eq:hdrda} is equivalent to
	\begin{align}
		D_{HDRDA}(\bm x) &= \argmin_k  (\bm x - \xbar_k)' \bm U_1 \bm W_k^{-1} \bm U_1' (\bm x - \xbar_k) + \ln | \bm W_k |. \label{eq:hdrda-decomposed}
	\end{align}

@@

Thus far, we have shown that the \emph{HDRDA} classifier in \eqref{eq:hdrda} is
invariant to the term $\bm U_2$, thus yielding an equivalent classifier in
\eqref{eq:hdrda-decomposed} with a substantial reduction in the computational
complexity. While the classifier is much improved in terms of computational
performance compared to the \emph{RDA} classifier originally proposed by
\cite{Friedman:1989tm}, we provide two additional results that further improve
the computational performance of our \emph{HDRDA} classifier. First, we
demonstrate that the inverse and determinants of $\bm W_k$ can be written so
that they are performed on $n_k \times n_k$ matrices rather than the larger $q
\times q$ matrices, resulting in further improvements to the computational
efficiency.

@@

In this section, we describe four high-dimensional microarray data sets and
compare our proposed \emph{HDRDA} classifier with five recently proposed
classifiers designed for small-sample, high-dimensional data: the \emph{DLDA}
classifier from \cite{Dudoit:2002ev}, the shrunken centroids \emph{RDA}
classifier from \cite{Guo:2007te}, the shrinkage-based \emph{DLDA} classifier
from \cite{Pang:2009ik}, the shrinkage-mean-based DLDA classifier from
\cite{Tong:2012hw}, and the penalized \emph{LDA} classifier from
\cite{Witten:2011kc}. The classifiers from \cite{Tong:2012hw} and
\cite{Pang:2009ik} are both modifications of the \emph{DLDA} classifier from
\cite{Dudoit:2002ev}, where the former employs an improved mean estimator and
the latter utilizes an improved variance estimator. Below, we refer to each
classifier by the first author's surname.

We evaluated the classification performance of each classifier by randomly
partitioning the data such that $2/3$ of the observations were allocated as
training data and the remaining $1/3$ of the observations were allocated as a
test data set. For each random partition, we applied the feature-selection
method from \cite{Dudoit:2002ev} to the training data to choose the $d$ features
having the largest ratio of their between-group to within-group sums of squares.
We then filtered the test data set with the corresponding $d$ features. After
training each classifier from the filtered training data, we classified the
filtered test data sets and computed the proportion of mislabeled test
observations to estimate the classification error rate for each
classifier. Repeating this process 1000 times, we computed the average of the
error-rate estimates for each classifier. We remark that our simulation design
avoids the selection bias discussed in \cite{Ambroise:2002fa}. We did not
include the \emph{RDA} classifier from \cite{Friedman:1989tm} in our studies to
avoid its excessive computations. In Table \ref{tab:data-summary}, we summarize
the attributes of each microarray data set.

@@

Next, we examine the optimal estimates of the tuning parameters for the
\emph{HDRDA} classifier. For brevity, we include only a graphical summary of the
optimal estimates of $\lambda$ for the Shipp data set in Figure
\ref{fig:optimal-shipp}. Similar figures for the other three data sets can be
found in the Supplementary Material. In Figure \ref{fig:optimal-shipp}, the
distribution of the optimal values of $\lambda$ selected through
cross-validation has a notable bathtub shape, resembling a Beta distribution
with both shape parameters less than unity. Specifically, the optimal values of
$\lambda$ are typically near 0 or near 1, corresponding to the \emph{QDA} and
\emph{LDA} classifiers, respectively. We noticed a similar distribution of
optimal values of $\lambda$ for the other three data sets.

@@

In our simulation study, we compared the \emph{HDRDA} classifier to several
recently proposed classifiers, including the Dudoit, Pang, Tong, and Guo
classifiers, each of which assumes independence of the features conditional on
class membership. Despite the rapid computational performance of diagonal
classifiers and the reduction in the number of parameters to estimate,
\cite{Mai:2012bf} and \cite{Fan:2012iq} have noted that diagonal classifiers
can often yield inferior classification performance to other classification
methods. We have shown that our proposed \emph{HDRDA} classifier often yields
superior classification accuracy to the diagonal classifiers considered,
confirming the assertions of \cite{Mai:2012bf} and
\cite{Fan:2012iq}. Furthermore, the \emph{HDRDA} classifier is able to compete
with the diagonal classifiers in terms of speed by incorporating a new
dimension-reduction technique.
